### 2018-09-15 source/Ch1-ThisThesis/intro.tex
added neal comments

\pg
We then outline the application of this work, along with other modern tools, to the creation of a large, deep and high-resolution image of a rich patch of sky. While this approach is potentially extremely rewarding in terms of the science one can do with its resulting maps, it is also uniquely complicated. There exist many LOFAR HBA ($\sim$120-240 MHz) sky surveys \citepads[cf.]{2016MNRAS.460.2385W,2017A&A...598A.104S}, but they do not yet make use of international LOFAR, and are thus limited in resolution ($\sim 5''$). And while there are surveys made at LOFAR frequencies ($\sim$150MHz) with other instruments \citepads[e.g.]{2017A&A...598A..78I}, they do not match the resolution attainable with international LOFAR. The aim of our image is then to show that the international LOFAR stations can be used to create a radio image that matches HST resolution - which would allow for significant cross-matching and source analysis, since there is HST data available for the field being imaged, the Extended Groth Strip (EGS). This would not be a true survey, but more of a proof of concept: the application of new tools and the development of an appropriate strategy to image a large patch of the sky with the full LOFAR array. This image would have better resolution than FIRST \citepads[see][and references therein]{2015ApJ...801...26H}, which has a $4''$ resolution, relatively low sensitivity, and no short baselines (thereby missing diffuse emission and extended structure) at $\sim$1.4 GHz, while still being sensitive to diffuse emission due to the use of the LOFAR core stations. 
%Finally, there will of course be the SKA1-MID survey once the SKA is operational; this survey is expected to provide excellent resolution, sensitivity, and good $uv$-coverage, making it sensitive to diffuse emission as well as point sources.

>>>>

\pg
We then outline the application of this work, along with other modern tools, to the creation of a large, deep and high-resolution image of a rich patch of sky. While this approach is potentially extremely rewarding in terms of the science one can do with its resulting maps, it is also uniquely complicated. There exist many LOFAR HBA ($\sim$120-240 MHz) sky surveys \citepads[cf.]{2016MNRAS.460.2385W,2017A&A...598A.104S}, but they do not yet make use of international LOFAR, and are thus limited in resolution ($\sim 5''$). And while there are surveys made at LOFAR frequencies ($\sim$150MHz) with other instruments \citepads[e.g.]{2017A&A...598A..78I}, they do not match the resolution attainable with international LOFAR. The aim of our image is then to show that the international LOFAR stations can be used to create a radio image that matches HST/ACS resolution - which would allow for significant cross-matching and source analysis, since there is HST data available for the field being imaged, the Extended Groth Strip (EGS). This would not be a true survey, but more of a proof of concept: the application of new tools and the development of an appropriate strategy to image a large patch of the sky with the full LOFAR array. This image would have better resolution than FIRST \citepads[see][and references therein]{2015ApJ...801...26H}, which has a $4''$ resolution, relatively low sensitivity, and no short baselines (thereby missing diffuse emission and extended structure) at $\sim$1.4 GHz, while still being sensitive to diffuse emission due to the use of the LOFAR core stations. 
%Finally, there will of course be the SKA1-MID survey once the SKA is operational; this survey is expected to provide excellent resolution, sensitivity, and good $uv$-coverage, making it sensitive to diffuse emission as well as point sources.

### 2018-09-15 source/Ch2-techniques/RIME.tex
added neal comments

\pg
The Radio Interferometer's Measurement Equation, or RIME, is a formalism which allows for an elegant and efficient formulation of the physical processes which affect signal propagation, from astrophysical effects to instrumental effects. Its fundamental underlying hypothesis is \emph{linearity}: that transformations along the signal are linear with respect to the basis chosen to represent our signal.

@@

\pg
For $\mathbf{e}$ to be a 2-vector, rather than a 3-vecotr, we must choose to use $xyz$ as our coordinates, with $z$ the direction of propagation of our signal. \citetads{1996A&AS..117..137H} then show that the following relation holds:
\begin{equation}
2 \evector \evector^\Herm = 2  \begin{bmatrix} <e_x e_x^\star> & <e_x e_y^\star> \\ <e_y e_x^\star> & <e_y e_y^\star> \end{bmatrix} = \begin{bmatrix} I+Q & U+iV \\ U-iV & I-Q \end{bmatrix} = \Bmatrix
\end{equation}

@@

\pg
Until this point, we have not considered any kind of sky variability in time. In effect, \cref{eq.fullsky.RIME} describes the snapshot taken by an interferometer for a single measurement. Limiting ourselves to this scenario, however, strongly limits the sampling of the $uv$-plane which can be performed. Indeed, interferometers generally rely on the Earth's rotation to rotate their baselines $pq$, thus increasing $uv$-coverage at no additional cost.This technique is known as \emph{super-synthesis} (\cite{supersynthesis}\footnote{\href{http://www.jpier.org/PIERB/pierb22/05.10032105.pdf}{(link to paper)}}) For \cref{eq.fullsky.RIME} to hold throughout an observation, we must therefore assume that $B_{app}$ remains constant in time over the course of the observation. In effect, this means assuming that
\begin{equation}\label{eq.selfcal.requirement}
\Ejones_p(t,\lvec) = \Ejones_p(\lvec) = \Ejones(\lvec) \quad \forall (t,p)
\end{equation}

>>>>

\pg
The Radio Interferometer's Measurement Equation, or RIME, is a formalism which allows for an elegant and efficient formulation of the physical processes which affect signal propagation, from astrophysical effects to instrumental effects. Its fundamental underlying hypothesis is \emph{linearity}: that transformations along the signal are linear with respect to the basis chosen to represent our signal. In other words, this means assuming that propagation effects are separable by antenna.

@@

\pg
For $\mathbf{e}$ to be a 2-vector, rather than a 3-vector, we must choose to use $xyz$ as our coordinates, with $z$ the direction of propagation of our signal. \citetads{1996A&AS..117..137H} then show that the following relation holds:
\begin{equation}
2 \evector \evector^\Herm = 2  \begin{bmatrix} <e_x e_x^\star> & <e_x e_y^\star> \\ <e_y e_x^\star> & <e_y e_y^\star> \end{bmatrix} = \begin{bmatrix} I+Q & U+iV \\ U-iV & I-Q \end{bmatrix} = \Bmatrix
\end{equation}

@@

\pg
Until this point, we have not considered any kind of sky variability in time. In effect, \cref{eq.fullsky.RIME} describes the snapshot taken by an interferometer for a single measurement. Limiting ourselves to this scenario, however, strongly limits the sampling of the $uv$-plane which can be performed. Indeed, interferometers generally rely on the Earth's rotation to rotate their baselines $pq$, thus increasing $uv$-coverage at no additional cost.This technique is known as \emph{super-synthesis} (\cite{supersynthesis}\footnote{\href{http://www.jpier.org/PIERB/pierb22/05.10032105.pdf}{\url{http://www.jpier.org/PIERB/pierb22/05.10032105.pdf}}}) For \cref{eq.fullsky.RIME} to hold throughout an observation, we must therefore assume that $B_{app}$ remains constant in time over the course of the observation. In effect, this means assuming that
\begin{equation}\label{eq.selfcal.requirement}
\Ejones_p(t,\lvec) = \Ejones_p(\lvec) = \Ejones(\lvec) \quad \forall (t,p)
\end{equation}

### 2018-09-15 source/Ch2-techniques/calibration.tex
added neal comments

\pg
In practice, of course, self-cal will be limited by noise. More precisely, it will be limited by sensitivity, in the form of the signal-to-noise ratio (henceforth SNR). For sufficiently bright sources, with high SNR, self-calibration can easily improve dynamic range by a factor of 10 (\cite{serendipitous}, p. 154) in a single iteration.

>>>>

\pg
In practice, of course, self-cal will be limited by noise. More precisely, it will be limited by two factors: sensitivity, in the form of the signal-to-noise ratio (henceforth SNR), and the atmosphere (as sufficient SNR is needed per coherence time, which is a function of atmosphere activity). For sufficiently bright sources, with high SNR, self-calibration can easily improve dynamic range by a factor of 10 (\cite{serendipitous}, p. 154) in a single iteration.

### 2018-09-15 source/Ch2-techniques/science.tex
added neal comments

\pg
Of course, in reality, we know that electron populations in astrophysical sources do not all share exactly the same energy when emitting. The effect of the electron energy distribution must therefore be taken into account. Because the population in question is energised through non-thermal processes, its energy distribution is not Maxwellian, and is generally written as a power law: $S(E) \propto E^\alpha \rightarrow S_\nu \propto \nu^{\alpha}$, where $\alpha$ is the object's spectral index.
The question now becomes: what $\alpha$ is appropriate for what kind of object? Does every source have the same $\alpha$ at all frequencies? Do all synchrotron sources have the same spectral index at a given frequency, and if not, what does this tell us?

@@

\pg
In the optically-thick regime, we know that the most efficient radiation mechanism is simple black-body radiation, and so the power spectrum at a given frequency can be defined in terms of an ``electron temperature". In the Rayleigh-Jeans approxmation (valid as we are explicitly in the low-frequency regime), this gives $S_\nu \propto \nu^{5/2}$, or $\alpha=5/2$. 
In the optically-thin regime, meanwhile, we recover a mixture of the underlying power-law distribution ($S_\nu \propto e^{-\nu}$) and black-body radiation (which has a frequency-dependent $\alpha$). Empirically, we generally observe $\alpha \sim 0.7$.

>>>>

\pg
Of course, in reality, we know that electron populations in astrophysical sources do not all share exactly the same energy when emitting. The effect of the electron energy distribution must therefore be taken into account. Because the population in question is energised through non-thermal processes, its energy distribution is not Maxwellian, and is generally written as a power law: $S(E) \propto E^\alpha \rightarrow S_\nu \propto \nu^{-\alpha}$, where $\alpha$ is the object's spectral index. Note that the index of the energy distribution (which relates the number of electrons at a given energy) is different from the synchrotron index, with the two related by the following equation: $\alpha=-(\gamma-1)/2$.
The question now becomes: what $\alpha$ is appropriate for what kind of object? Does every source have the same $\alpha$ at all frequencies? Do all synchrotron sources have the same spectral index at a given frequency, and if not, what does this tell us?

@@

\pg
In the optically-thick regime, we know that the most efficient radiation mechanism is simple black-body radiation, and so the power spectrum at a given frequency can be defined in terms of an ``electron temperature". In the Rayleigh-Jeans approxmation (valid as we are explicitly in the low-frequency regime), this gives $S_\nu \propto \nu^{5/2}$, or $\alpha=5/2$. 
In the optically-thin regime, meanwhile, we recover a mixture of the underlying power-law distribution ($S_\nu \propto \nu^{-\alpha}$) and black-body radiation (which has a frequency-dependent $\alpha$). Empirically, we generally observe $\alpha \sim 0.7$.

### 2018-09-15 source/Ch4-Qweights/paper.tex
final commit for sake of backup - thesis completed
added neal comments

\section{Discussion}

@@

>>>>

\clearpage
\section{Discussion}

@@

\pg
Once the conditioning issues of the artefact-optimal weights are resolved, a Briggs-type implementation of the overall adaptive quality-based weighting scheme with a sliding parameter could conceivably be implemented. As things stand, however, there is no good reason to use the currently-existing artefact-optimal weighting scheme over the well-conditioned sensitivity-optimal weighting scheme.

### 2018-09-15 source/Ch7-3C295/3c295.tex
added neal comments

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes compensate for this effect, and minimise the associated image deterioration. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602 of SB155}
\end{figure}

@@

\pg
So far, we have shown the gain solutions for one subband, out of a total of 324 used subbands. The spectral behaviour of the amplitude and phase solutions can give us insight into whether we are absorbing true flux or calibration artefacts into our model: if our calibration model becomes more accurate to the true underlying flux, we expect e.g. the phase of core station calibration solutions to remain the same (as we are improving high-resolution details of our model) while our international station calibration solutions would become flatter in amplitude. Similarly, we expect to see large-scale frequency-dependent behaviour in the phase solutions of international stations, with smaller fluctuations around this behaviour: the effect of differential TEC, ionospheric turbulence, etc. Indeed, while the core stations (and, to a lesser extent, the remote stations) can be approximated as having the ``same ionosphere" above them, this is not the case for international stations: this is a major source of the difficulties in calibrating these stations. As such, for international stations, we expect to see complex structure in the phase gain solutions, characteristic of the differences between the local ionosphere above the core and the international stations. For core stations, we only expect to see structure characteristic of scintillation (regular stripes with no frequency-dependent structure) in the phase gain solutions.

@@

\pg
We see that some of the structure in the amplitude gains shown in \cref{fig.fullBW.DE601HBA} goes away over the course of self-calibration, though one of the two arcs present in the plot appears to decrease in amplitude. For \cref{fig.fullBW.DE602HBA}, we see that the pattern in the amplitude plot clearly changes - the fact that the pattern in \cref{fig.fullBW.DE602HBA.before} does not appear to be either dispersion or scintillation pattern would indicate that it consists of unmodeled flux. While some structure remains in the last 2 hours and low frequencies for this antenna's gain solution amplitudes, the last 2 hours of the observation are generally poor compared to the first six: we expect that the quality-based weighting scheme will minimise the impact of these poor calibration solutions on the final images. In \cref{fig.fullBW.DE603HBA}, we see that the self-calibration process was clearly less effective than it was for DE602HBA: while some of the structure was minimised (in the first two hours particularly: we go from the peculiar ``ellipse" amplitude pattern to something which looks more like scintillation), a lot clearly remains. Our self-calibration process could therefore likely be taken further. However, in our last step, we use 324 out of the full 366 subbands (leaving some out due to poor quality); a single round of self-calibration with this full dataset takes 5000 minutes on the best computing cluster available for this work. Because a single round of self-calibration would, at this point, take as much time as imaging a portion of the EGS due to computational constraints, we treat the self-calibration process as ``having converged" at this point: while it could be yet improved, we have reached a signal-to-noise ratio of 2.78 mJy.

>>>>

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes were found to compensate for this effect quite well, and minimise the associated image deterioration\footnote{The evidence for this is in the previous chapter, as the test data is the same as used here; nevertheless, an extra test was carried out in the context of these calibration and imaging tests for the sake of completeness, and found that quality-based weighting schemes resulted in better results than simply flagging the last 2 hours.}. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602 of SB155}
\end{figure}

@@

\pg
So far, we have shown the gain solutions for one subband, out of a total of 324 used subbands. The spectral behaviour of the amplitude and phase solutions can give us insight into whether we are absorbing true flux or calibration artefacts into our model: if our calibration model becomes more accurate to the true underlying flux, we expect e.g. the phase of core station calibration solutions to remain the same (as we are improving high-resolution details of our model) while our international station calibration solutions would become flatter in amplitude. Similarly, we expect to see large-scale frequency-dependent behaviour in the phase solutions of international stations, with smaller fluctuations around this behaviour: the effect of differential TEC, ionospheric turbulence, etc. Indeed, while the core stations (and, to a lesser extent, the remote stations) can be approximated as having the ``same ionosphere" above them, this is not the case for international stations: this is a major source of the difficulties in calibrating these stations. As such, for international stations, we expect to see complex structure in the phase gain solutions, characteristic of the differences between the local ionosphere above the core and the international stations. For core stations, we only expect to see structure characteristic of scintillation (regular stripes with no frequency-dependent structure) in the phase gain solutions\footnote{This approach was taken due to time constraints: verifying that our model was compatible with the closure phases and amplitudes would have been a better test to perform, but would also likely have taken much longer to be fruitfully interpreted.}.

@@

\pg
We see that some of the structure in the amplitude gains shown in \cref{fig.fullBW.DE601HBA} goes away over the course of self-calibration, though one of the two arcs present in the plot appears to decrease in amplitude. For \cref{fig.fullBW.DE602HBA}, we see that the pattern in the amplitude plot clearly changes - the fact that the pattern in \cref{fig.fullBW.DE602HBA.before} does not appear to be either dispersion or scintillation pattern would indicate that it consists of unmodeled flux. While some structure remains in the last 2 hours and low frequencies for this antenna's gain solution amplitudes, the last 2 hours of the observation are generally poor compared to the first six: we expect that the quality-based weighting scheme will minimise the impact of these poor calibration solutions on the final images. In \cref{fig.fullBW.DE603HBA}, we see that the self-calibration process was clearly less effective than it was for DE602HBA: while some of the structure was minimised (in the first two hours particularly: we go from the peculiar ``ellipse" amplitude pattern to something which looks more like scintillation), a lot clearly remains. Our self-calibration process could therefore likely be taken further. However, in our last step, we use 324 out of the full 366 subbands (leaving some out due to poor quality); a single round of self-calibration with this full dataset takes 5000 minutes on the best computing cluster available for this work. Because a single round of self-calibration would, at this point, take as much time as imaging a portion of the EGS due to computational constraints, we treat the self-calibration process as ``having converged" at this point: while it could be yet improved, we have reached a noise of 2.78 mJy.
